---
title: "Assignment 9: Class Wrap-up"
format: html
editor: visual
author: "Amber Camp"
date: "`r Sys.Date()`"
---

## Assignment 9: Class Wrap-up

We are finally at the end of the semester! The final assignment for the class will be to summarize and share some of the things we've done so far.

The plan is for everyone to be randomly paired with a tiny topic pertaining to something we've learned together this semester and then submit a code snippet to be presented in class on Wednesday, 11/12. (Think \~2 minute walkthrough)

#### Libraries

```{r, message=FALSE}
library(tidyverse)
```

#### Read in the list of topics

```{r, message=FALSE}
# topics <- read_csv("data/cs201_topics.csv")
# view(topics)

# anything to edit?
# topics[7, "topic"] <- "wildcard"
```

#### Read in the list of student names

```{r, message=FALSE}
# names <- read_csv("data/cs201_names.csv")
# view(names)
```

#### Generate a random number in the names dataframe

```{r, message=FALSE}
# names <- names %>%
#   mutate(number = sample(1:nrow(topics), nrow(names), replace = FALSE))
# 
# view(names)
```

#### Join the names to the topics

```{r, message=FALSE}
# assigned_topics <- names %>%
#   full_join(topics, by = "number")
# 
# print(assigned_topics)
# 
# write.csv(assigned_topics, "data/assignment9_assigned_topics.csv")

# note that if you run the above script and randomize the names again, your outcome will be different from mine!!
# go with the topics assigned in class and announced on Canvas!
```

### Instructions

1.  Find your assigned topic. Talk to me if you reeeeeeaaaally want to do something else, or trade with someone (but let me know).

2.  Prepare a code snippet that briefly (\~2) explains the topic in the code chunk below, using `penguins` data.

3.  Push to GitHub **BY 10:00am** on Wednesday 11/10 (I'll need time to compile before class).

4.  Come to class Wednesday and show us how it's done :)

#### Extra libraries needed (if any)

```{r, message=FALSE}
library(tidyverse) # tidyverse is a collection that includes tidyr, dplyr, ggplot2, and more!
library(here)
library(palmerpenguins)
library(janitor)
library(plotly)
```

#### Read in penguins data

```{r, message=FALSE}
penguins <- read_csv(here("data/penguins_lter.csv"))

penguins <- penguins %>% 
  clean_names()
```

#### customizing plots (labs, themes, titles)

#### labs/titles: adds titles, subtitles, and axis labels.

```{r, message=FALSE}
ggplot(data = penguins, aes(x = flipper_length_mm, fill = species)) +
  geom_histogram(alpha = 0.5) +
  theme_classic() +
  scale_fill_manual(values = c("darkorange","darkorchid","cyan4")) +
  labs(title = "Palmer Penguins, Distribution of Flipper Length",
       subtitle = "3 penguin species observed on three islands in the Palmer Archipelago, Antarctica",
       x = "Flipper Length (mm)",
       y = "Number of Penguins")
```

#### themes: applies a black and white theme or background. There is other themes like theme_minimal(), classic() and more.

```{r}
ggplot(data = penguins, aes(x = flipper_length_mm, fill = species)) +
  geom_histogram(alpha = 0.5, bins = 30) +
  labs(
    title = "Palmer Penguins: Distribution of Flipper Length",
    subtitle = "Three species observed in the Palmer Archipelago, Antarctica",
    x = "Flipper Length (mm)",
    y = "Number of Penguins"
  ) +
  theme_linedraw()
```

#### fill: scale_fill_manual(): manually pick colors for values. In the penguins case its species and it matches the penguins species category. You can change the colors to whatever color you want.

```{r}
ggplot(data = penguins, aes(x = flipper_length_mm, fill = species)) +
  geom_histogram(alpha = 0.5, bins = 30) +
  theme_dark() +
  scale_fill_manual(values = c("pink", "yellow", "cyan")) +
  labs(
    title = "Palmer Penguins: Distribution of Flipper Length",
    subtitle = "Three species observed in the Palmer Archipelago, Antarctica",
    x = "Flipper Length (mm)",
    y = "Number of Penguins"
  )
```
